---
title: "Tryggve.query1.v2"
author: "Lu Yi, Kaarina Kowalec, Patrick Sullivan"
date: "05/05/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: textmate
    df_print: paged
    code_folding: hide
params:
  arg1: "/Users/luyi/Downloads/tryggve"
  arg2: "NOR"
  arg3: " "
  arg4: " "
  arg5: " "
  arg6: " "
---

<style type="text/css">
  .main-container {
    max-width: 1500px !important;
  }
  /* Whole document: */
  body{
    font-family: Helvetica;
    font-size: 12pt;
  }
</style>



***
## Prep step 
#### Setting up R pkgs
- Perform sanity check if the R pkgs exist   
- Install the missing packages if any

```{r setup, message =FALSE, warning= FALSE}
# add ",message =FALSE, warning= FALSE" to suppress messages. 

#==== check if the required R packages are installed
#     & install the package(s) if not there
required.Rpackages <- 
  c("tidyverse", "data.table", "kableExtra", "knitr", "summarytools")

new.packages <- required.Rpackages[!(required.Rpackages %in% installed.packages()[,"Package"])]

if (length(new.packages)) { 
  # if there are new packages, install
  install.packages(new.packages) 
  } else {
  # if all required packages are there, just load 
library(tidyverse)
library(data.table)
library(kableExtra)

library(knitr)
# plus some suggested settings to improve the html output
opts_chunk$set(results = 'asis',      # This is essential (can also be set at the chunk-level)
              comment = NA, 
              prompt = FALSE, 
              cache = FALSE)

library(summarytools)
# plus some suggested settings to improve the html output
st_options(plain.ascii = FALSE,       # This is very handy in all Rmd documents
           style = "rmarkdown",       # This too
           footnote = NA,             # Avoids footnotes which would clutter the results
           subtitle.emphasis = FALSE  # this might improve the headings' layout
           )    
}
```

```{r, echo=FALSE}
st_css()   # separate this option to avoid a long text of messages
```


#### Read parameters and data
- Perform sanity check if the master file exists  
  -  No? - print error message and stop execution  
  -  Yes! - continue to load the files  

<p style="color:blue">*Check the output below - look out for Error or Success message.* </p>  

```{r read all pheno}
#=== setup variables
MAIN_DIR  <- params$arg1
COUNTRY <- params$arg2
trait <-params$arg3

contVar=params$arg5

#MasterFile_name <- paste0(COUNTRY, ".tryggve.master.file.tsv")
#MasterFile_path <- paste(MAIN_DIR, MasterFile_name, sep="/")

MasterFile_path <- params$arg4
# sanity check if the master file exists

if (file.exists(MasterFile_path)) {
  print("Successfully locate the master file, continue processing Query 1")
} else {
  print("ERROR: master file does not exist, stop processing Query 1");
  quit()
}

#=== get master file
Master <- fread(MasterFile_path, stringsAsFactors = FALSE, data.table = FALSE)

#===  read in all pheno data
# remove the dataset final if it already exists
if (exists("final") == TRUE) { rm(final) }
# loop over the files in the master list
for(i in 1:nrow(Master)) {
  a <- Master %>% slice(i:i)
  b <- paste(a$tcb_folder, a$tcb_pheno, sep="/")
  c <- fread(b, stringsAsFactors = FALSE, data.table = FALSE)
  d <- cbind(a, c) %>% select(-tcb_folder, -tcb_plink, -tcb_pheno)
  if (exists("final") == FALSE) { final <- d } else { final <- bind_rows(final, d) }
  rm(a,b,c,d)
}

str(final)
```
<p style="color:blue">*Check if the data have been read in properly.*</p>  
*e.g., for toydata, the dataframe 'final' should have 8 vars and 600 obs* 

***
**Next - process the queries** 

- Query 1.1 - are all tcb_fid-iid combinations in final unique?  
- Query 1.2 - summarise data for all samples  
- Query 1.3 - summarise the 3 key variables, for each tcb_cohort-tcb_subcohort  


## Query 1.1 - check unique IDs
### Table 1. Number of unique IDs in cohort-subcohort
```{r query 1.1}

# Check the combination of IDs within each cohort-subcohort 
query1.1 <- final %>% 
  # concatenate the FID-IID
  mutate(fid_iid = paste(IID,IID, sep="_")) %>%
  # for each cohort-subcohort combination
  group_by(tcb_cohort, tcb_subcohort) %>%
  # calculate number of rows, and number of unique FID-IID
  summarise(
    N_row=n(),
    N_uniqID=length(unique(fid_iid))
    ) %>% 
  # add an indicator showing whether all IDs are unique
  mutate(I_alluniqIDs = ifelse(N_uniqID == N_row, 1, 0))

# display nicely
query1.1 %>% 
  # use kable() to make a nice table
  kable() %>% 
  kable_styling("striped", full_width = F) 

# save the query1.1 results 
query1.1_name <- paste0(COUNTRY, ".tryggve.query1.1.tsv")
query1.1_path <- paste(MAIN_DIR, query1.1_name, sep="/")

fwrite(query1.1, file = query1.1_path, quote = F, sep="\t")

# print out which data subsets contains duplicate IDs, if any
for (i in c(1:nrow(query1.1))) {
  if (query1.1$I_alluniqIDs[i] == 1) {i=i+1} 
  else { print(query1.1[i,]) }
}

# N.B. no output here means no duplicate IDs. 
```
<p style="color:blue">*Check in the table - 'I_alluniqIDs' == 0 indicates duplicate IDs. * </p>  



***
## Query 1.2 - summarise data for all samples
### Table 2. Descriptive statistics across the entire cohort
```{r description tables for all samples, results='asis'}

df <- final %>% 
  # drop id variables when describing data
  select(-IID) %>% 
  # set cohort and subcohort as factor variables 
  mutate(
    tcb_cohort = as.factor(tcb_cohort),
    tcb_subcohort = as.factor(tcb_subcohort)
  )
# print nice summary tables using dfSummary()
dfSummary(df, 
          plain.ascii = FALSE, style = "grid", 
          graph.magnif = 0.75, valid.col = FALSE, headings=F)


```
<p style="color:blue">*Check if variable distributions here and onwards roughly agree to what's known about this cohort.*</p>  




## Query 1.3 - summarise the 3 key variables, for each tcb_cohort-tcb_subcohort



### Table 3. Sex by cohort-subcohort
```{r sex}
# Discrete variable: Sex


n_samples <- df %>% group_by(tcb_cohort,tcb_subcohort) %>% summarise(N_row=n())

n_sex <- df %>% group_by(tcb_cohort,tcb_subcohort,Sex) %>% summarise(N_sex=n())

query1.3_sex <- full_join(n_samples, n_sex, by=c("tcb_cohort","tcb_subcohort")) %>% 
  mutate(prop_sex = (N_sex/N_row)*100 ) %>% 
  ungroup()

# save the descriptive data on sex 
query1.3_sex_name <- paste0(COUNTRY, ".tryggve.query1.3_sex.tsv")
query1.3_sex_path <- paste(MAIN_DIR, query1.3_sex_name, sep="/")

fwrite(query1.3_sex, file = query1.3_sex_path, quote = F, sep="\t")

# display nicely
query1.3_sex %>% 
  pivot_wider(names_from = Sex, values_from = c(N_sex, prop_sex),values_fill = list(seen = 0)) %>% 
  select(tcb_cohort, tcb_subcohort, N_sex_F, prop_sex_F, N_sex_M, prop_sex_M) %>% 
  # rename for better presentation, together with add_header_above() row below
  rename(N_F = N_sex_F, "%_F" = prop_sex_F,
         N_M = N_sex_M, "%_M" = prop_sex_M) %>%
  # use kable() to make a nice table
  kable() %>% 
  kable_styling("striped", full_width = F) %>%
  # uncommnent the following if want group rows
#  pack_rows("Cohort1", 1, 3) %>%
#  pack_rows("Cohort2", 4, 6) %>%
  add_header_above(c(" ", " ", "Female" = 2, "Male" = 2))



```




### Table 4. Interested trait by cohort-subcohort
```{r mdd}
# Discrete variable: AnyF33
v<-toString(params$arg3)


traitA<-as.name(v)

cat("interested trait:")
print(traitA)

n_mdd <- df %>% group_by(tcb_cohort,tcb_subcohort, {{traitA}} ) %>% summarise(N_mdd=n())

query1.3_mdd <- full_join(n_samples, n_mdd, by=c("tcb_cohort","tcb_subcohort")) %>% 
  mutate(prop_mdd = (N_mdd/N_row)*100 ) %>%
  ungroup()

# save the descriptive data on mdd 
query1.3_mdd_name <- paste0(COUNTRY, ".tryggve.query1.3_mdd.tsv")
query1.3_mdd_path <- paste(MAIN_DIR, query1.3_mdd_name, sep="/")

fwrite(query1.3_mdd, file = query1.3_mdd_path, quote = F, sep="\t")


# display nicely
query1.3_mdd %>% 
  pivot_wider(names_from = {{traitA}} , values_from = c(N_mdd, prop_mdd),values_fill = list(seen = 0)) %>% 
  select(tcb_cohort, tcb_subcohort, N_mdd_1, prop_mdd_1, N_mdd_0, prop_mdd_0) %>% 
  # rename for better presentation, together with add_header_above() row below
  rename(N_CASE = N_mdd_1, "%_CASE" = prop_mdd_1,
         N_CONT = N_mdd_0, "%_CONT" = prop_mdd_0) %>%
  # use kable() to make a nice table
  kable() %>% 
  kable_styling("striped", full_width = F) %>%
  # uncommnent the following if want group rows
 # pack_rows("Cohort1", 1, 3) %>%
#pack_rows("Cohort2", 4, 6) %>%
  add_header_above(c(" ", " ", "MDD_TRUE" = 2, "MDD_FALSE" = 2))

```


### Table 5. Continious variates by cohort-subcohort
```{r fig.cap="This is a caption" , results='asis'}



vv=params$arg5

v2=unlist(strsplit(vv, split=","))




query1.3_pca <-""
a=''
b=''
for(i in 1:length(v2)) {   #The counter declaration
  #then do something for each iteration

  v<-toString(v2[i])

  str1='continuous_grouped_descr3'
  str2=v
  gr_name=paste(str1,str2,sep="")
  
  gr_name2=as.name(gr_name) 

# Putting description of continuous variable (pc) into one table
 gr_name2 <- stby(data = get(v,df), 
                      INDICES = df[, c("tcb_cohort","tcb_subcohort")], 
                     FUN = descr, stats = "common", round.digits = 0)
# the table below still displays many digits...

st_pc=paste('query1.3_pc',v,sep="")
st_name=paste('query1.3_pc_name',v,sep="")
#st_path=paste('query1.3_pc_path',v,sep="")

st_fl=paste("tryggve.query1.3_pc.tsv",v,".tsv",sep="")

st_pc2=toString(st_pc)
st_name2=toString(st_name)


query1.3_pc=as.name(st_pc)
query1.3_pc_name=as.name(st_name)

dummy1  <- gr_name2 %>% tb() %>% arrange(tcb_cohort)
assign(st_pc2, dummy1)
dummy2 <-paste0(COUNTRY, st_fl)
assign(st_name2,dummy2)
#query1.3_pc_name <- paste0(COUNTRY, ".tryggve.query1.3_pc.tsv")
query1.3_pc_path <- paste(MAIN_DIR, as.name(st_name2) , sep="/")
  





fwrite(assign(st_pc2, dummy1), file = query1.3_pc_path, quote = F, sep="\t")

# in order to print in a loop we need to use following structure, name change in each loop was not necessary.

a <- assign(st_pc2, dummy1) %>%   select(-variable) 



print(knitr::kable(head(a),  caption = paste0 ("Statistical properties of: ", v) , full_width = F,  table.attr = "class=\"striped\"",  format = "html",  align = "lccrr")  %>% kable_styling("striped", full_width = F)  )  

cat('\n\n<!-- -->\n\n')



}



```



### Table 6. Binary variables by cohort-subcohort
```{r Binary variables}
# Discrete variable: AnyF33





vv=params$arg6

v2=unlist(strsplit(vv, split=","))

for(i in 1:length(v2)) {  

  
v<-toString(v2[i])  

if (v!='Sex') {

traitA<-as.name(v)

n_mdd <- df %>% group_by(tcb_cohort,tcb_subcohort, {{traitA}} ) %>% summarise(N_mdd=n())

query1.3_mdd <- full_join(n_samples, n_mdd, by=c("tcb_cohort","tcb_subcohort")) %>% 
  mutate(prop_mdd = (N_mdd/N_row)*100 ) %>%
  ungroup()

# save the descriptive data on mdd 
query1.3_mdd_name <- paste0(COUNTRY, ".tryggve.query1.3_mdd.tsv")
query1.3_mdd_path <- paste(MAIN_DIR, query1.3_mdd_name, sep="/")

fwrite(query1.3_mdd, file = query1.3_mdd_path, quote = F, sep="\t")



# display nicely
a <- query1.3_mdd %>% 
  pivot_wider(names_from = {{traitA}} , values_from = c(N_mdd, prop_mdd),values_fill = list(seen = 0)) %>% 
  select(tcb_cohort, tcb_subcohort, N_mdd_1, prop_mdd_1, N_mdd_0, prop_mdd_0) %>% 
  # rename for better presentation, together with add_header_above() row below
  rename(N_CASE = N_mdd_1, "%_CASE" = prop_mdd_1,
         N_CONT = N_mdd_0, "%_CONT" = prop_mdd_0) 
  # use kable() to make a nice table
#  kable() %>% 
#  kable_styling("striped", full_width = F) %>%
  # uncommnent the following if want group rows
 # pack_rows("Cohort1", 1, 3) %>%
#pack_rows("Cohort2", 4, 6) %>%
#  add_header_above(c(" ", " ", "MDD_TRUE" = 2, "MDD_FALSE" = 2))

print(knitr::kable(head(a),  caption = paste0 ("Binary Trait: ", v) , full_width = F,  table.attr = "class=\"striped\"",  format = "html",  align = "lccrr")  %>% kable_styling("striped", full_width = F)  )  

cat('\n\n<!-- -->\n\n')
}
}
```



## Timestamp
```{bash}
# echo "The program was completed by <INSERT ANALYST NAME HERE> at `date`"
echo "The program was completed by Lu Yi at `date`"
```

