---
title: "Tryggve.query1.v2"
author: "Lu Yi, Kaarina Kowalec, Patrick Sullivan"
date: "05/05/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: textmate
    df_print: paged
    code_folding: hide
params:
  arg1: "/Users/luyi/Downloads/tryggve"
  arg2: "NOR"

---

<style type="text/css">
  .main-container {
    max-width: 1500px !important;
  }
  /* Whole document: */
  body{
    font-family: Helvetica;
    font-size: 12pt;
  }
</style>



### Purpose
This script aims to create a utility program which does the followings:  

+ process desired queries  
+ output metadata on the basic pheno  

<p style="color:red">**N.B. the output from these queries contains aggregated values ONLY - *no inidividual-level data are written out or displayed*.**</p>

***
## Instructions


#### Parameters:
1. path of the main directory  
2. country code  

#### Input files:
1. genotype master file    
2. phenotype file  

#### Required file format:  

**Input #1: geno master file**  
*This is a space- or tab-delimited tsv file which contains country or cohort names, file path and file names. One master file per country.*  
An example is provided in example.tryggve.master.file.tsv, and displayed below.
```{r}
if (file.exists("example.tryggve.master.file.tsv")) {
  tmp1 <- read.table("example.tryggve.master.file.tsv", header=T)
  head(tmp1)
} else {
  print("ERROR: example file missing, ask Lu Yi")
}

```
  
**Input #2: pheno file**  
*This is a space- or tab-delimited tsv file which contains geno-matched IDs and basic phenotypes. One pheno file per cohort.*  
The 4 key variables are:  

- IID
- 2 string variable, Sex ("M"/"F") and MDD (1/0)  
- 1 continuous variable, tcb_age  

A toy example is provided in example.tryggve.pheno.file.tsv, and displayed below (not real personal data).
```{r}
if (file.exists("example.tryggve.pheno.file.tsv")) {
  tmp2 <- read.table("example.tryggve.pheno.file.tsv", header=T)
  head(tmp2)
} else {
  print("ERROR: example file missing, ask Lu Yi")
}

```

